---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rwpfunctionality

<!-- badges: start -->
<!-- badges: end -->

This package offers users comprehensive documentation on surveys conducted to assess water point functionality and management committees. These surveys were carried out across nine countries in both sub-Saharan Africa and South Asia.

## Installation

You can install the development version of rwpfunctionality from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("openwashdata/rwpfunctionality")
```

## Project goal

Sustained functionality of rural water systems is a global challenge. Addressing premature failures and unreliable performance of water points is crucial to achieve the 2030 Sustainable Development Goal (SDG) 6, which calls for universal and equitable access to safe and affordable drinking water for all (WHO, 2017). As of 2020, 771 million people lacked a basic water service (WHO & UNICEF, 2021). Yet even for those with water services, unreliable water infrastructure that provides intermittent supply, or is broken down or abandoned, contributes to unsustainable and unsafe water access worldwide (Allaire et al., 2018; Burt et al., 2018; Klug et al., 2018; Valcourt et al., 2020).


## Data

The data set includes observations of water point functionality and management committee questionnaires.

The package provides access to one single data set.

```{r, echo = TRUE}
library(rwpfunctionality)
```

The `rwpfunctionality` data set has `r ncol(rwpfunctionality)`
variables and `r nrow(rwpfunctionality)` observations. For an overview
of the variable names, see the following table.

```{r, eval=FALSE}
rwpfunctionality
```

```{r, echo=FALSE}
readr::read_csv("data-raw/dictionary.csv", col_types = cols(.default = "c")) |> 
  dplyr::filter(file_name == "rwpfunctionality.rda") |> 
  dplyr::select(variable_name:description) |> 
  knitr::kable()
```

## Example

### ...

This is a basic example which shows you how to solve a common problem:

```{r example}
library(rwpfunctionality)
## basic example code
```

## License

Data are available as 
[CC-BY](https://github.com/openwashdata/rwpfunctionality/LICENSE.md).

## Citation

To cite this package, please use:

```{r}
citation("rwpfunctionality")
```


