---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rwpfunctionality

<!-- badges: start -->
<!-- badges: end -->

The goal of rwpfunctionality is to ...

## Installation

You can install the development version of rwpfunctionality from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("openwashdata/rwpfunctionality")
```

## Project goal
The primary goal of BaSIS (Basic Sanitation Information System) within the context of implementing Community-Led Total Sanitation (CLTS) in Ghana is to combat open defecation. By collecting data on communities practicing open defecation, the project aims to shift away from previous subsidy-based approaches and focus on community engagement. BaSIS seeks to empower communities, particularly in regions like Upper West, Eastern, Northern, Volta, and Central, to abandon open defecation practices through CLTS. The objective is to break the cycle of fecal-oral contamination, reduce the spread of diseases linked to poor sanitation, and promote sustainable, community-driven solutions to improve overall hygiene and sanitation practices.


## Data

The data set includes the locations of the existing open defecation free (ODF) communities in 
various districts of Ghana. It was last updated in September 2017. 

The package provides access to one data set.

```{r, echo = TRUE}
library(rwpfunctionality)
```

The `rwpfunctionality` data set has `r ncol(rwpfunctionality)`
variables and `r nrow(rwpfunctionality)` observations. For an overview
of the variable names, see the following table.

```{r, eval=FALSE}
rwpfunctionality


## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(rwpfunctionality)
## basic example code
```

## License

Data are available as 
[CC-BY](https://github.com/openwashdata/rwpfunctionality/LICENSE.md).

## Citation

To cite this package, please use:

```{r}
citation("rwpfunctionality")
```


